---
title: "p8105_hw1_yy3297"
author: "Yunxi Yang"
date: "2022-09-25"
output: "html_document"
---

# Problem 1

## setup and load the data set

```{r setup}
library(tidyverse)
data("penguins", package = "palmerpenguins")
```

## Description of the penguins data set

1.  The Penguins data set includes 8 variables, they are `r colnames(penguins)`.
2.  This data set has `r nrow(penguins)` rows, and `r ncol(penguins)` columns.
3.  The mean flipper length of this data set is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`.

## Plot a scatter plot of flipper_length_mm (y) vs bill_length_mm (x)

```{r plot a scatterplot of flipper_length_mm (y) vs bill_length_mm (x)}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color=species)) + 
  geom_point() + 
  ggtitle('Scatterplot of flipper_length_mm (y) vs bill_length_mm') +
  xlab('Bill Length (mm)') +
  ylab('Flipper Length (mm)')
```

## Export the scatterplot to project directory
```{r export scatterplot to project directory}
ggsave("Flipper_vs_Bill_Scatterplot_Penguins.pdf")
```

# Problem 2

## Create a data frame according to p2 requirements
```{r create a data frame}
p2_df = tibble(
  nrom = rnorm(10, mean = 0, sd = 1),
  logi_vector = nrom > 0,
  char_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  fac_vector = factor(c("High", "Middle", "Low", "High", "High", "Middle", "Low", "Low", "Midlle", "Middle"))
)
view(p2_df)
```

## Calculate the mean of variables in the dataframe
```{r calculate the mean of each variable}
nrom_mean = mean(pull(p2_df, var = nrom))
nrom_mean
logi_vector_mean = mean(pull(p2_df, var = logi_vector))
logi_vector_mean
char_vector_mean = mean(pull(p2_df, var = char_vector))
char_vector_mean
fac_vector_mean = mean(pull(p2_df, var = fac_vector))
fac_vector_mean
```
Conclusion: The _mean_ function works on the nrom and logical vectors. However, it does not work on the character vectors and factor vectors.

## Apply as.numeric function to logical, character, and factor variables
```{r}
logi_vector_mean2 = mean(as.numeric(pull(p2_df, var = logi_vector)))
logi_vector_mean2
char_vector_mean2 = as.numeric(pull(p2_df, var = char_vector))
char_vector_mean2
fac_vector_mean2 = as.numeric(pull(p2_df, var = fac_vector))
fac_vector_mean2
```
Conclusion: The as.numeric function applies successfully to logical and factor vector, while fails to apply to character vectors. This indicates that characters cannot be converted to numbers, while logical vectors and factors can be converted to numbers for further calculations.


